---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# flint

<!-- badges: start -->
[![R-CMD-check](https://github.com/etiennebacher/flint/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/etiennebacher/flint/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


`flint` is a small R package to find and replace lints in R code.

* Lints detection with `lint()`
* Automatic replacement of lints with `fix()`
* Compatibility with (some) `{lintr}` rules
* Fast
* Low-dependency

`flint` is powered by [`astgrepr`](https://github.com/etiennebacher/astgrepr/),
which is itself built on the Rust crate [`ast-grep`](https://ast-grep.github.io/).

## Installation

### Windows or macOS

``` r
install.packages(
  'flint', 
  repos = c('https://etiennebacher.r-universe.dev', getOption("repos"))
)
```

### Linux

``` r
install.packages(
  'flint', 
  repos = c('https://etiennebacher.r-universe.dev/bin/linux/jammy/4.3', getOption("repos"))
)
```

## Usage

Optional setup:

* `setup_flint()`: creates the folder `flint` and populates it with built-in rules
  as well as a cache file. You can modify those rules or add new ones if you
  want more control.
  
You can use `flint` as-is, without any setup. However, running `setup_flint()`
enables the use of caching, meaning that the subsequent runs will be faster. It
is also gives you a place where you can store custom rules for your 
project/package.
  
The everyday usage consists of two functions:

* `lint()` looks for lints in R files;
* `fix()` looks for lints in R files and automatically applies their replacement 
  (if any).
  
One can also experiment with `flint::lint_text()` and `flint::fix_text()`:

```{r}
flint::lint_text("
any(is.na(x))
any(duplicated(y))
")
flint::fix_text("
any(is.na(x))
any(duplicated(y))
")
```


## Comparison with existing tools

The most used tool for lints detection in R is `lintr`. However, `lintr`'s 
performance is not optimal when it is applied on medium to large packages. Also,
`lintr` cannot perform automatic replacement of lints. 

`styler` is a package to clean code by fixing indentation and other things, but 
doesn't perform code replacement based on lints. 

`flint` is extremely performant. This is a small benchmark on 3.5k lines of 
code with only three linters:

```{r}
file <- system.file("bench/test.R", package = "flint")

bench::mark(
  lintr = lintr::lint(
    file, linters = list(lintr::any_duplicated_linter(), lintr::any_is_na_linter(),
                         lintr::matrix_apply_linter())
  ),
  flint = flint::lint(
    file, linters = list(flint::any_duplicated_linter(), flint::any_is_na_linter(),
                         flint::matrix_apply_linter()), 
    open = FALSE
  ),
  check = FALSE
)
```

## Contributing

Did you find some bugs or some errors in the documentation? Do you want
`flint` to support more rules?

Take a look at the [contributing
guide](https://flint.etiennebacher.com/CONTRIBUTING.html) for
instructions on bug report and pull requests.

## Acknowledgements

The website theme was heavily inspired by Matthew Kay’s `ggblend`
package: <https://mjskay.github.io/ggblend/>.
