---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tinylint

`tinylint` is a small R package to detect lints in R code and optionally
replace them.

* Lint detections with `lint()`
* Automatic replacement of lints with `fix()`
* Compatibility with `{lintr}` rules
* Extremely fast
* Low-dependency

`tinylint` is powered by [`astgrepr`](https://github.com/etiennebacher/astgrepr/),
which is itself built on the Rust crate [`ast-grep`](https://ast-grep.github.io/).


## Usage

Start by setting up `tinylint` with `tinylint::setup_tinylint()`. This imports
`sgconfig.yml` and stores a set of rules in `inst/tinylint/rules`. 

`tinylint` provides two families of functions: 

* those for linting: `lint()`, `lint_text()`. 
* those for replacing lints: `fix()`, `fix_text()`


## Comparison with existing tools

The most used tool for lints detection in R is `lintr`. However, `lintr`'s 
performance is not optimal when it is applied on medium to large packages. Also,
`lintr` cannot perform automatic replacement of lints. 

`styler` is a package to clean code by fixing indentation and other things, but 
doesn't perform code replacement based on lints. 

`tinylint` is extremely performant. This is a small benchmark on 3.5k lines of 
code with only three linters:

```{r}
library(bench)
library(lintr)
library(tinylint, warn.conflicts = FALSE)

file <- system.file("bench/test.R", package = "tinylint")

bench::mark(
  lintr::lint(file, linters = list(any_duplicated_linter(), any_is_na_linter(), matrix_apply_linter())),
  tinylint::lint(file, linters = c("any_duplicated", "any_na", "matrix_apply"), open = FALSE),
  check = FALSE
)
```

One can also experiment with `tinylint::lint_text()`:

```{r}
lint_text("any(duplicated(x))")
```

