---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# flint

`flint` is a small R package to find and replace lints in R code.

* Lints detection with `lint()`
* Automatic replacement of lints with `fix()`
* Compatibility with (some) `{lintr}` rules
* Fast
* Low-dependency

`flint` is powered by [`astgrepr`](https://github.com/etiennebacher/astgrepr/),
which is itself built on the Rust crate [`ast-grep`](https://ast-grep.github.io/).


## Usage

Optional setup:

* `setup_flint()`: creates the folder `flint` and populates it with built-in rules
  as well as a cache file. You can modify those rules or add new ones if you
  want more control.
  
You can use `flint` as-is, without any setup. However, running `setup_flint()`
enables the use of caching, meaning that the subsequent runs will be faster. It
is also the place where you can store custom rules.
  
The everyday usage consists of two functions:

* `lint()` looks for lints in R files;
* `fix()` looks for lints in R files and automatically applies their replacement 
  (if any).
  
One can also experiment with `flint::lint_text()` and `flint::fix_text()`:

```{r}
flint::lint_text("
any(is.na(x))
any(duplicated(y))
")
flint::fix_text("
any(is.na(x))
any(duplicated(y))
")
```


## Comparison with existing tools

The most used tool for lints detection in R is `lintr`. However, `lintr`'s 
performance is not optimal when it is applied on medium to large packages. Also,
`lintr` cannot perform automatic replacement of lints. 

`styler` is a package to clean code by fixing indentation and other things, but 
doesn't perform code replacement based on lints. 

`flint` is extremely performant. This is a small benchmark on 3.5k lines of 
code with only three linters:

```{r}
file <- system.file("bench/test.R", package = "flint")

bench::mark(
  lintr = lintr::lint(
    file, linters = list(lintr::any_duplicated_linter(), lintr::any_is_na_linter(),
                         lintr::matrix_apply_linter())
  ),
  flint = flint::lint(
    file, linters = list(flint::any_duplicated_linter(), flint::any_is_na_linter(),
                         flint::matrix_apply_linter()), 
    open = FALSE
  ),
  check = FALSE
)
```
