---
title: "Adding new rules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding new rules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(flint, warn.conflicts = FALSE)
```

`flint` comes with an extensive set of rules taken from `lintr`, but if 
necessary one can also extend it relatively easily. This will require some
knowledge of the Rust crate `ast-grep` to write. This crate
has great documentation on [creating new rules](https://ast-grep.github.io/guide/pattern-syntax.html)
so you should start there.

In this vignette, our objective is to replace calls to `stop()` by 
`rlang::abort()`, for instance because we prefer the formatting of the output
with the latter.
```{r, error=TRUE}
stop("this is an error")

rlang::abort("this is an error")
```

## Step 1: setup

First, we need to set up `flint` using `setup_flint()`. This will create a 
`flint` folder that contains (among other things) a `rules` folder where all 
rules are stored. This is divided between `builtin` rules (that shouldn't modified manually) and `custom` rules (where we will store our custom rule).

Then, we can create the structure of a new rule by copying an existing rule, say
`flint/rules/builtin/any_is_na.yml` for instance. Removing the stuff that is
specific to this rule, we end up with this structure:

```yaml
id: 
language: r
severity: warning
rule:
  pattern: 
fix: 
message: 
```


## Step 2: start exploring

When we use `stop()`, we can pass multiple elements, like in `paste0()`:

```{r, error=TRUE}
n <- 10
stop("Got ", n, " values instead of 1.")
```

This is not possible with `rlang::abort()`, which needs everything to be in the
argument `message`, meaning that we need to manually put all those elements in 
`paste0()`:

```{r, error=TRUE}
n <- 10
rlang::abort(paste0("Got ", n, " values instead of 1."))
```

Therefore, we can look for the pattern `stop(...)` and replace it by 
`rlang::abort(paste0(...))`. Capturing all elements in a pattern is done with
`$$$` and those elements can be used in the `fix` or `message` arguments using 
by wrapping them in `~~`:

```yaml
id: stop_abort-1
language: r
severity: warning
rule:
  pattern: stop($$$ELEMS)
fix: rlang::abort(paste0(~~ELEMS~~))
message: Use `rlang::abort()` instead of `stop()`.
```

After storing this rule in `flint/rules/custom/stop_abort.yml`, we can
call:

```{r, eval=FALSE}
flint::lint_text(
  'stop("Got ", n, " values instead of 1.")', 
  linters = "stop_abort"
)
```

Running our example with `lint_text()` now shows the message:

```{.r}
flint::lint_text(
  'stop("Got ", n, " values instead of 1.")', 
  linters = "stop_abort"
)
#> Original code: stop("Got ", n, " values instead of 1.") 
#> Suggestion: Use `rlang::abort()` instead of `stop()`. 
#> Rule ID: stop_abort-1 
```
And `fix_text()` correctly applies the fix:

```{r, eval=FALSE}
flint::fix_text(
  'stop("Got ", n, " values instead of 1.")', 
  linters = "stop_abort"
)
#> Old code: stop("Got ", n, " values instead of 1.") 
#> New code: rlang::abort(paste0("Got ", n, " values instead of 1.")) 
```

Note that there are still some corner cases to address, such as ignoring the
arguments `call.` and `domain` of `stop()` if they are specified, but this is
not in the scope of this vignette.


# Step 3 (optional) 

To automatically use this new linter without having to specify it manually,
we can add it to `flint/config.yaml`:

```yaml
keep:
  - any_duplicated
  [...]
  - unreachable_code
  - stop_abort
```

Running `lint_text()` or `fix_text()` without `linters` now works:

```{r, eval=FALSE}
flint::fix_text('stop("Got ", n, " values instead of 1.")')
#> Old code: stop("Got ", n, " values instead of 1.") 
#> New code: rlang::abort(paste0("Got ", n, " values instead of 1.")) 
```

# Step 4

The new linter is now set up, you can use `flint` as before to lint or fix
specific files or entire folders, e.g. `flint::fix_dir("R")`.
